---
title: "Shakespeare Play Analysis"
output:
  pdf_document
documentclass: article
classoption: a4paper  
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Import libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidytext)
library(readr)
library(stringr)
```

## Import and Tidy Data

The purpose of the sample project is describing the overall atmosphere of the prominent plays of Shakespeare. We will be analyzing words from the plays to conclude whether his masterpieces are either sentimentally positive or negative depending on the type such as Comedy or Tragedy. Hypothetically, Comedy should have more positive atmosphere and vice versa.
First, we will count how many titles and types are in Shakespeare.

```{r, message = FALSE, warning = FALSE}
# Import Six Shakespeare Acts' Data
shakespeare = read_csv("shakespeare.csv")

# Pipe the shakespeare data frame to the next line
shakespeare %>% 
  # Use count to find out how many titles/types there are
  count(title,type) %>%
  arrange(desc(n))
```

*Hamlet, Prince of Denmark*, seems to have the most lines compared to other five Shakespeare Plays, having **6776 lines**.
Next, we will redistribute texts into words and respective line numbers.

```{r}
tidy_shakespeare <- shakespeare %>%
  # Group by the titles of the plays
  group_by(title) %>%
  # Define a new column linenumber
  mutate(linenumber = row_number()) %>%
  # Transform the non-tidy text data to tidy text data
  unnest_tokens(word, text) %>%
  ungroup() %>%
  select(-X1)
head(tidy_shakespeare)
```

\break

## Plot the Cleaned Shakespeare Data

Next, we are going to use **tidy_shakespeare** to analyze Six Shakespeare plays. Which how does each sentiment fluctuate from the beginning to the end?

```{r, message= FALSE, fig.align = "center"}
# Plot Sentiment Changes with Respect to PLays
tidy_shakespeare %>%
  inner_join(get_sentiments("bing")) %>%
  count(title, type, index = linenumber %/% 70, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
         fixed_title = str_wrap(title, width = 20)) %>%
  # Put index on x-axis, sentiment on y-axis, and map comedy/tragedy to fill
  ggplot(aes(x = index, y = sentiment, fill = type)) +
  # Make a bar chart with geom_col()
  geom_col() +
  # Separate panels for each title with facet_wrap()
  facet_wrap(~ fixed_title, scales = "free_x") +
  labs(x = "Index of Shakespeare Plays by 70 Lines Each",
       y = "Sentiment of Each Plays") +
  scale_fill_discrete(name = "Type")
```

## Conclusion

As expected, comedic Shakespeare plays tend to have more positive sentiments while tragic acts tend to be more negative.
Notice that *The Tragedy of Romeo and Juliet* follows the somewhat positive introduction and then the tragic end.